---
title: "Deductible Comparison"
output:
  html_document: default
  pdf_document: default
header-includes: 
  - \usepackage{subfig}
  - \captionsetup[figure]{labelformat=empty}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this exercise is to look at a comparison of auto insurance deductibles and their costs against a range of anticipated repair costs. 

To jump to the plots, click [here](#plot)

## Libraries

For this comparison we will use:

1. `dplyr` for data exploration/manipulation
2. `ggplot2` for data visualization
3. `grid` and `gridExtra` for document formatting

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
```
## Data
Next we will create our dataset by manually establishing deductibles and their annual cost (as vectors), and use the `seq` function to generate a sequence of anticipated repair costs. This range will be joined using the `merge` function to create a table of deductibles, costs and repair costs.

Finally, we will use `dplyr`'s mutate function to calculate the repair cost covered after the deductible and a comparison of the cost covered to the annual cost of the deductible.
```{r data}
deductible <- c(100,250,500,1000,1500,2500,5000)

deductible_costs <- 
  data.frame(
    deductible_factor = 
      factor(
        scales::dollar(deductible), 
        ordered = TRUE,
        levels = scales::dollar(deductible)[order(deductible)]
      ),
    deductible = deductible,
    deductible_cost = c(1262,1193,1093,948,830,652,371)
  ) %>%
  merge(
    .,
    data.frame(repair_cost = seq(from=100,to=10000,by=100))
  ) %>%
  mutate(
    repair_cost_covered = pmax(repair_cost-deductible,0),
    deductible_cost_vs_covered_cost = repair_cost_covered/deductible_cost,
    monthly_deductible_cost = deductible_cost/12
  )
```

## Plot

With the data prepared, we will use `ggplot` to plot the result.

```{r plot_render, anchor="figure", fig.show='hold', out.width="50%",fig.cap="Figure 1: Plot and table comparing deductibles and costs.", fig.subcap=c("Plot", "Table")}
test <- ggplot(
  deductible_costs %>% filter(repair_cost < 5000), 
  aes(x = repair_cost, y = deductible_cost_vs_covered_cost)
) + 
  geom_line(aes(color = deductible_factor)) +
  labs(
    x = "Repair Cost ($)",
    y = "Repair Cost Covered vs Deductible Cost ($)",
    color = "Deductible"
  ) +
  scale_colour_brewer(palette = "Set1") +
  scale_x_continuous(labels = scales::dollar_format(prefix="$")) +
  scale_y_continuous(
    labels = function(x){paste0(x*100, "%")},
    breaks = round(
      seq(
        min(deductible_costs$deductible_cost_vs_covered_cost),
        max(deductible_costs$deductible_cost_vs_covered_cost),
        by = 1
      )
    )
  )

test2 <- as.data.frame(
  deductible_costs %>% 
    group_by(deductible_factor) %>% 
    summarise(
      cost = scales::dollar(max(deductible_cost)),
      monthly_cost = scales::dollar(max(monthly_deductible_cost))
    )
)

test

grid.newpage()
grid.table(d=test2, rows=NULL, cols=c("Deductible", "Annual Cost", "Monthly Cost"), theme=ttheme_minimal())
```